---
title: "Scout transition of pure spores"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---
# Overview
Heat-treated a wild-type Bacillus culture (KBS0812) to kill of vegetative cells. 
Then followed the number of vegetative and spores colonies after. 
Aim was to see if trends in vegetative abundance which would be consistnent with germination
This could affect mutation rate (compared to if there's no turnover)

# Set working directory and load packages
```{r}
rm(list = ls())
getwd()
setwd("~/GitHub/SporMut")
require("bbmle")
require("grid")
require("png")
```

# Load data
```{r}
# load raw data
data <- read.table("data/spore.scout.txt", header = TRUE)
t0 <- data$date[1]
days <- as.numeric(difftime(data$date,t0, units="days"))
hours <- days *24
data <- data.frame(days,hours,data[2:6])

# calculate total CFUs using dilution 
# multiply be 10 to go from 100 ul dilution to 1 mL
tot.ab <- data$total * 10^(data$total_dil) *10

# calculate spore CFUs using dilution (per 1000 uL dilution)
# multiply be 10 to go from 100 ul dilution to 1 mL
spo.ab <- data$spore * 10^(data$spor_dil) *10

# calculate veg CFUs using dilution
veg.ab <- tot.ab - spo.ab

# calculate proportion of vegetative CFUs
prop.veg <- 1 - (spo.ab / tot.ab)

# calculate proportion of spore CFUs
prop.spor <- 1 - prop.veg

# make new data frame
scout <- data.frame(data, tot.ab, veg.ab, spo.ab, prop.veg, prop.spor)

scout.A <- scout[ which(scout$rep == 'A'), ]
scout.B <- scout[ which(scout$rep == 'B'), ]
scout.C <- scout[ which(scout$rep == 'C'), ]
scout.D <- scout[ which(scout$rep == 'D'), ]

#scout <- scout[which(scout$prop.veg > 0), ]

# Is percent vegetative cells greater than zero?
t.test(scout$prop.veg, mu=0)





# write table
write.table(scout, "data/spore.scout.output.txt", sep = "\t",
              col.names = T, row.names = T)
```

# Fit data with sigmoidal function using MLE
```{r, results='hide', warning=FALSE, message=FALSE}
#
# # Initial parameters
# A = 0.95 # Maximum
# B = 0.11 # Minimum
# M = 9 # Midpoint
# W = -21 # Slope
# Z = 2 # Error
#
# # Function
# fit <- mle2(prop.veg ~ dnorm(mean = b + (a - b) / (1 + exp((m - time.d) / w)), sd = z),
#             start = list(a = A, b = B, m = M, w = W, z = Z), data = trans)
```

# Plot data and add sigmoidal function
```{r}
png(filename="~/GitHub/SporMut/output/scout-veg-percent.png",
     width = 900, height = 900, res = 96*2)

par(mar = c(5, 7, 5, 7))

plot(scout$day, scout$prop.veg, pch = 21, 
          bg = "grey", col = "grey", lwd = 2, cex.axis = 1.25, xlab = NA, ylab = NA,
          xlim = c(-5, 60), ylim = c(-0.6, 1.1), yaxt = "n", xaxt = "n")
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = expression(-0.5, 0.0, 0.5, 1.0),
     at = c(-0.5, 0, 0.5, 1.0))

axis(side = 4, lwd.ticks = 2, cex.axis = 1.25,
    at=c(-0.5, 0, 0.5, 1.0), labels = F)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25,
    at = c(0, 25, 50), labels = T)

axis(side = 3, lwd.ticks = 2, cex.axis = 1.25,
    at = c(0, 25, 50), labels = F)

mtext('Time (d)', side = 1, outer = TRUE, cex = 1.5, 
      line = -2, adj = 0.5)

mtext('Proportion vegetative', side = 2, outer = TRUE, cex = 1.5, 
      line = -3.5, adj = 0.5)

# curve(coef(fit)[[2]] + (coef(fit)[[1]] - coef(fit)[[2]]) /
#        (1 + exp((coef(fit)[[3]] - x) / coef(fit)[[4]])),
#        from = 0, to = 240, add = TRUE, lty = 2, lwd = 2.5, col = "black")

#0812A
regA.per <- lm(scout.A$prop.veg ~ scout.A$days)
clip(0.05,60,-0.6,1)
A.int <- regA.per$coefficients[1]
A.slp <- regA.per$coefficients[2]
abline(a = A.int, b = A.slp, col = "red", lwd = 2.5, lty = 6)

#0812B
regB.per <- lm(scout.B$prop.veg ~ scout.B$days)
clip(0.05,60,-0.6,1)
B.int <- regB.per$coefficients[1]
B.slp <- regB.per$coefficients[2]
abline(a = B.int, b = B.slp, col = "red", lwd = 2.5, lty = 6)

#0812C
regC.per <- lm(scout.C$prop.veg ~ scout.C$days)
clip(0.05,60,-0.6,1)
C.int <- regC.per$coefficients[1]
C.slp <- regC.per$coefficients[2]
abline(a = C.int, b = C.slp, col = "red", lwd = 2.5, lty = 6)

#0812D
regD.per <- lm(scout.D$prop.veg ~ scout.D$days)
clip(0.05,60,-0.6,1)
D.int <- regD.per$coefficients[1]
D.slp <- regD.per$coefficients[2]
abline(a = D.int, b = D.slp, col = "red", lwd = 2.5, lty = 6)

dev.off()
graphics.off()

img <- readPNG("~/GitHub/SporMut/output/scout-veg-percent.png")
grid.raster(img)
```

###STOPPED HERE

# Plot population densisites over time 
```{r}
png(filename="~/GitHub/SporMut/figures/pop.densities.png",
     width = 1200, height = 900, res = 96*2)

par(mar = c(5, 7, 5, 7))

plot(trans$time.d, log10(trans$veg.ab), pch = 21, 
          bg = "grey80", cex = 1.25, col = "grey80", lwd = 0.75, cex.axis = 1.25, xlab = NA, ylab = NA,
          xlim = c(-25, 250), ylim = c(1.5, 9), yaxt = "n", xaxt = "n")
points(trans$time.d, log10(trans$spo.ab), pch = 21, lwd = 0.75, col = "grey50", bg = "grey50", cex = 1.25)

# https://stackoverflow.com/questions/19612348/break-x-axis-in-r
# xgap <- ifelse(trans$time.d > 1500, trans$time.d-1295, trans$time.d)
# plot(xgap, log10(trans$veg.ab), pch = 21, 
#           bg = "red", cex = 1.25, col = "black", lwd = 1, cex.axis = 1.25, xlab = NA, ylab = NA,
#           xlim = c(-25, 300), ylim = c(1.5, 9), yaxt = "n", xaxt = "n")
# points(xgap, log10(trans$spo.ab), pch = 22, lwd = 1, bg = "blue", cex = 1.25)

legend(145, 8.75, c("Spore", "Non-spore"), pch = 21, pt.bg = c("grey50", "grey80"), col = c("grey50", "grey80"),
       pt.cex = 1.25, pt.lwd = 0.75, bty = 'n', y.intersp = 0.80, lwd = 0,
       x.intersp = -0.5)

box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = expression(10^2, 10^4, 10^6, 10^8),
     at = c(2, 4, 6, 8))

axis(side = 4, lwd.ticks = 2, cex.axis = 1.25,
    at=c(2, 4, 6, 8), labels = F)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25,
    at = c(0, 50, 100, 150, 200, 250), labels = T)

axis(side = 3, lwd.ticks = 2, cex.axis = 1.25,
    at = c(0, 50, 100, 150, 200, 250), labels = F)

mtext('Time (d)', side = 1, outer = TRUE, cex = 2, 
      line = -2, adj = 0.5)

mtext(expression('Bacteria (CFU ml'^-1*')'), side = 2, outer = TRUE, cex = 2, 
      line = -3.5, adj = 0.5)

#mtext('Spores', side = 1, line = -8.5, at = 265, cex = 0.75)
#mtext('Non-spores', side = 1, line = -7.5, at = 280, cex = 0.75)

# Fit non-spore abundance with sigmoidal function using MLE
# Initial parameters
A = 6 # Maximum 
B = 6 # Minimum
M = 5 # Midpoint
W = -21 # Slope
Z = 2 # Error

# Function
fit2 <- mle2(log10(veg.ab) ~ dnorm(mean = b + (a - b) / (1 + exp((m - time.d) / w)), sd = z),
            start = list(a = A, b = B, m = M, w = W, z = Z), data = trans) 

curve(coef(fit2)[[2]] + (coef(fit2)[[1]] - coef(fit2)[[2]]) / 
        (1 + exp((coef(fit2)[[3]] - x) / coef(fit2)[[4]])), 
      from = 0, to = 235, add = TRUE, lty = 2, lwd = 2.5, col = "black") 

# Fit spore abundance with sigmoidal function using MLE
# Initial parameters
A = 6 # Maximum 
B = 6 # Minimum
M = 5 # Midpoint
W = 21 # Slope
Z = 2 # Error

# Function
fit3 <- mle2(log10(spo.ab) ~ dnorm(mean = b + (a - b) / (1 + exp((m - time.d) / w)), sd = z),
            start = list(a = A, b = B, m = M, w = W, z = Z), data = trans) 

curve(coef(fit3)[[2]] + (coef(fit3)[[1]] - coef(fit3)[[2]]) / 
        (1 + exp((coef(fit3)[[3]] - x) / coef(fit3)[[4]])), 
      from = 0, to = 235, add = TRUE, lty = 2, lwd = 2.5, col = "black") 

dev.off()
graphics.off()

# Show Plot
img <- readPNG("figures/pop.densities.png")
grid.raster(img)
```
